---
title: "Fatalities from the 2021 Military Coup in Myanmar. Data from AAPP (Burma)"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: [ "twitter", "facebook", "menu" ]
    theme: flatly
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)

library(flexdashboard)
```

Visualisations
===================================== 


```{r import-and-prepare-data}
library(tidyverse)
library(cowplot)
library(plotly) # remotes::install_github("ropensci/plotly")

# from https://aappb.org/
pdf_file <- list.files(pattern = ".pdf")[1]

library("tabulizer")
tab <- extract_tables(pdf_file, method = "lattice")
# head(tab[[1]])

# clean the data 
library("janitor")
library("lubridate")
aappb_data_tbl <- 
map(tab, as_tibble) %>% 
  bind_rows() %>% 
  row_to_names(row_number = 1) %>% 
  clean_names() %>% 
  filter(!str_detect(tolower(name), "total")) %>% 
  mutate(age = parse_number(age)) %>% 
  mutate(date = str_extract(deceased_date, ".*21")) %>% 
  mutate(date = dmy(date)) %>% 
  mutate(Township = str_squish(str_remove(township, 
                                          "Tsp|Twnship|Township|township|T/s|Townsip|twonship"))) %>% 
  mutate(Township = str_replace(Township, "N/", "North ")) %>% 
  mutate(multiple_people_one_row = parse_number(name)) %>% 
  slice(rep(seq_len(n()), ifelse(is.na(multiple_people_one_row), 1, multiple_people_one_row)))

# timeline, cumulative and per day
aappb_data_tbl_date_n <- 
  aappb_data_tbl %>% 
  group_by(date) %>% 
  tally %>% 
  drop_na() %>% 
  arrange(date) %>% 
  mutate(
    `fatalities this day` = n,
    `total fatalities to date` = cumsum(n)
    )
```


Column {data-width=200}
-----------------------------------------------------------------------

### Cumulative Total of Fatalities

```{r}
# cumulative
people_killed_so_far <- 
  ggplot(aappb_data_tbl_date_n) +
    aes(date, `total fatalities to date`) +
    geom_line() +
    labs(y = paste0("People killed since\n", 
                    min(aappb_data_tbl_date_n$date)),
         x = "") +
    theme_minimal(base_size = 14)

# ggplotly(people_killed_so_far)

aappb_data_tbl_date_n_xts <- 
aappb_data_tbl_date_n %>% 
  dplyr::select(date,
                `total fatalities to date`) 

aappb_data_tbl_date_n_xts_2 <- 
  timetk::tk_xts(aappb_data_tbl_date_n_xts,
                 date_col = "date")
  

library(dygraphs)
dygraph(aappb_data_tbl_date_n_xts_2, 
        ylab = paste0("People killed since\n", 
                    min(aappb_data_tbl_date_n$date)))
```

### Fatalities per day

```{r}
# per day
people_killed_per_day <- 
  ggplot(aappb_data_tbl_date_n) +
    aes(date, `fatalities this day` ) +
    geom_col() +
    labs(y = "People killed\nper day",
         x = "") +
    theme_minimal(base_size = 14)

ggplotly(people_killed_per_day)
```

Column {data-width=200}
-----------------------------------------------------------------------

### Myanmar map: Fatalities per Township

```{r}
# Map, shapefile from https://hub.arcgis.com/datasets/ae5cb1d5972e429dba6ff415f5538961_0
library(sf)

township_map <- 
  st_read("4e3f01a4-6376-4cbb-a14f-54ba66960b7a2020329-1-133w8zn.1sut.shp",
          quiet = TRUE) %>% 
  mutate(Township = as.character(TS)) 

# check for variations in district names
# district_map %>% 
#  pull(District) %>% sort

# check for variations in township names
# township_map %>%  pull(Township) %>% sort

aappb_data_tbl_tsp_names_tally <- 
  aappb_data_tbl %>% 
  group_by(Township) %>% 
  tally %>% 
  filter(Township != "") %>% 
  mutate(Township = case_when(
    #            in aaappd data ~ in ArcGIS data
    Township == "Aungmyaytharz an" ~ "Aungmyaythazan",
    Township == "Kyauk Si" ~ "Kyaukse",
    Township == "Shwe Pyi Thar" ~ "Shwepyithar",
    Township == "Taungtwin Gyi" ~ "Thandaunggyi",
    Township == "North Okkalapy" ~ "North Okkalapa",
    Township == "Myingyin" ~ "Myingyan",
    Township == "Pwint Phyu" ~ "Pwintbyu",
    Township == "Zayyarthiri" ~ "Zay Yar Thi Ri",
    Township == "East Dagon" ~ "Dagon Myothit (East)",
    Township == "North Dagon" ~ "Dagon Myothit (North)",
    Township == "South Dagon" ~ "Dagon Myothit (South)",
    Township == "Kalay" ~ "Kale",
    Township == "Mandalay" ~ "Aungmyaythazan",
    Township == "Hlaing Tharyar" ~ "Hlaingtharya",
    TRUE ~ as.character(Township)
  )) 

# check names with ArcGIS data
# aappb_data_tbl_tsp_names_tally %>% pull(Township) %>% sort
  
township_map_kill_data <- 
township_map %>% 
  left_join(aappb_data_tbl_tsp_names_tally) %>% 
  mutate(map_label = paste0(Township, " (n = ", n, ")"))

# how many matches? 
# sum(township_map_kill_data$n, na.rm = TRUE)
```


```{r eval = FALSE}
# map with sf
# to handle some issue with multiple polys with same value
# https://github.com/ropensci/plotly/issues/1641
p_2 <- ggplot(township_map_kill_data) +
  geom_sf(aes(fill = n, 
              text = map_label, 
              colour = FID),
          size = 0.1
          ) +
    scale_fill_viridis_c(na.value = "white")  +
  guides(colour = FALSE) +
  coord_sf(expand = TRUE,
             xlim = c(91.8, 101.3),
             ylim = c(9.3, 29),
           clip = "on") +
  theme_void()

gg_2 <- ggplotly(p_2,  
                 tooltip="text",
                 width = 400, 
                 height = 700 )

m <- list(
  l = 0,
  r = 0,
  b = 0,
  t = 0,
  pad = 1
)

gg_2 %>%
  style(
    hoveron = "fills",
    # override the color mapping
    line.color = toRGB("gray40"),
    # don't apply these style rules to the first trace, 
    # which is the background graticule/grid
    traces = seq.int(2, length(gg_2$x$data))
  ) %>%
  hide_legend() %>% 
  hide_colorbar() %>% 
  layout(autosize = F,
         xaxis = list(range = c(92, 101.3)),
         yaxis = list(range = c(9.3, 28.75)),
         margin = m)
```

```{r}
#---------
# map with leaflet
library(leaflet)

township_map_kill_data_lft <- 
township_map_kill_data %>% 
   st_centroid() %>% 
            as("Spatial") 

township_map_kill_data_lft_coords <- raster::geom(township_map_kill_data_lft)

township_map_kill_data_lft$lng <- township_map_kill_data_lft_coords[ , 2]
township_map_kill_data_lft$lat <- township_map_kill_data_lft_coords[ , 3]

township_map_kill_data_lft_n <- 
  township_map_kill_data_lft %>% 
  sf::st_as_sf() %>% 
  filter(!is.na(n))
  
  leaflet(township_map_kill_data_lft) %>% 
    setView(lng = 96, 
            lat = 20, 
            zoom = 5) %>% 
    addTiles() %>% 
    addCircleMarkers(lng = township_map_kill_data_lft_n$lng,
                     lat = township_map_kill_data_lft_n$lat,
                     radius = township_map_kill_data_lft_n$n,
                     color = "red",
                     stroke = FALSE,
                     popup = ~(township_map_kill_data_lft_n$map_label),
                     fillOpacity = 0.6)
   
```


Column {data-width=200}
-----------------------------------------------------------------------

### Demographics of fatalities

```{r, fig.height=5, fig.width=10}
# age and sex of the deceased
aappb_data_tbl_sex <- 
aappb_data_tbl %>% 
  group_by(sex) %>% 
  tally %>% 
  filter(sex %in% c("F", "M"))

# by sex
people_killed_by_sex <- 
  ggplot(aappb_data_tbl_sex) +
  aes(sex, n) +
  geom_col() +
  labs(y = "People killed",
       x = "Sex of deceased") +
  theme_bw()

# by age
people_killed_by_age <- 
  ggplot(aappb_data_tbl) +
  aes(age) +
  geom_histogram() +
  labs(y = "People killed",
       x = "Age at death in years") +
  theme_minimal(base_size = 16)


ggdraw(people_killed_by_age) +
  draw_plot(people_killed_by_sex,
            # x, y
            .6, .6, 
            # width, height
            .35, .35) 
```

### Fatalities by Township & Wordcloud on Remarks for Each Fatality 

```{r, fig.height=10, fig.width=10}
fatalities_by_township <- 
ggplot(
  aappb_data_tbl_tsp_names_tally 
  ) +
  aes(reorder(Township, n, FUN = sum),
      n) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Fatalities") +
   theme_bw(base_size = 16) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) 

# word cloud on remarks
library(quanteda)

set.seed(10)

dfm_remarks <- 
aappb_data_tbl %>% 
  pull(remarks) %>% 
  tolower %>% 
  dfm(remove = c(stopwords("english")),
      remove_punct = TRUE,
      remove_numbers = TRUE) %>%
   dfm_trim(min_termfreq = 3) 
  
  
  # basic wordcloud
wc <- ~{
  textplot_wordcloud(dfm_remarks, 
                   color = rev(RColorBrewer::brewer.pal(10, "RdBu")))
}

ggdraw(fatalities_by_township) +
  draw_plot(plot_grid(wc),
                # x, y
            .1, -0.1, 
            # width, height
            1.2, 1.2) 
```

Data sources
===================================== 

Column {data-width=200}
-----------------------------------------------------------------------

### The data

The data come from the [Assistance Association for Political Prisoners (Burma)](https://aappb.org/) ([Wikipedia](https://en.wikipedia.org/wiki/Assistance_Association_for_Political_Prisoners)) and their fatality lists posted with their 'daily briefing' [blog posts](https://aappb.org/?cat=7). They also post summaries to [Twitter](https://twitter.com/aapp_burma).

### Journalists & Observers

- [Poppy McPherson](https://twitter.com/poppymcp)
- [Hnin Zaw](https://twitter.com/hninyadanazaw)
- [Shoon Naing](https://twitter.com/Shoon_Naing)
- [Mratt Kyaw Thu](https://twitter.com/mrattkthu)
- [Jonathan Head](https://twitter.com/pakhead)
- [Melissa Crouch](https://melissacrouch.com/)
- [Thant Myint-U](https://twitter.com/thantmyintu)
- [Matthew Tostevin](https://twitter.com/TostevinM)
- [Soe Zeya Tun](https://twitter.com/soezeya)
- [Aye Min Thant](https://twitter.com/the_ayeminthant)
- [Cape Diamond](https://twitter.com/cape_diamond)

Column {data-width=200}
-----------------------------------------------------------------------

### Independent Media

Independent Myanmar Media:

- [Frontier Myanmar](https://twitter.com/FrontierMM)
- [Myanmar Now](https://twitter.com/Myanmar_Now_Eng)
- [The Myanmar Times](https://twitter.com/TheMyanmarTimes)

Independent International Media

- Daily summaries by [Aye Min Thant](https://twitter.com/the_ayeminthant) on [hkcnews](https://www.hkcnews.com/triangles/blogger/%20%20Aye%20Min%20Thant-430)


### Social Media

- [#WhatsHappeningInMyanmar](https://twitter.com/hashtag/WhatsHappeningInMyanmar)
- [#MilkTeaAlliance](https://twitter.com/hashtag/MilkTeaAlliance)

Column {data-width=200}
-----------------------------------------------------------------------






